### 2018-11-05 GenusOne.tex

\section{Computation in genus one}
We can apply the results from the previous sections, in particular theorems \ref{thm:main}
and \ref{thm:Hbms*as*ggrep}, to compute the groups $H_*(\gone^m)$ explicitly.

Recall that $\gone$ is generated by two Dehn twists $\tu$ and $\tv$ about two simple
closed curves $\u,\v\subset\Sigma_{1,1}$ that intersect once transversely; a presentation
for $\gone$ is given by
\[
 \gone=\left<\tu,\tv\, |\, \tu\tv\tu=\tv\tu\tv \right>
\]
and therefore $\gone$ is isomorphic to the braid group on three strands $\beta_3=\beta_3(\D)$.