### 2018-11-05 MAIN.tex

\documentclass{amsart}
\usepackage{amsmath,amssymb,amsthm,amsfonts,amscd,mathrsfs,mathtools}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[hyphens]{url}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{tikz-cd}

\hyphenation{e-qui-va-riant se-cond ho-mo-lo-gy co-ho-mo-lo-gy}

\theoremstyle{plain}
    \newtheorem{thm}{Theorem}[section]
    \newtheorem{lem}[thm]   {Lemma}
    \newtheorem{cor}[thm]   {Corollary}

\theoremstyle{definition}
    \newtheorem{defn}[thm]  {Definition}
    \newtheorem{conj}[thm]{Conjecture}

\input{macros.tex}

\begin{document}

\title{Splitting of the homology of the punctured mapping class group}

\author{Andrea Bianchi}

\address{Mathematics Institute,
University of Bonn,
Endenicher Allee 60, Bonn,
Germany
}

\email{bianchi@math.uni-bonn.de}

\date{\today}

%\keywords{Topological complexity, aspherical spaces, Lusternik-Schnirelmann category, cohomological dimension, topological robotics, braid groups}
%\subjclass[2010]{55M99, 55P20 (Primary); 55M30, 20J06, 68T40 (Secondary).}

\begin{abstract}
Let $\ggm$ be the mapping class group of an orientable surface $\sgm$ of genus $g$ with one parametrised
boundary component and $m$ permutable punctures, and let $\bms$ be the braid group on $m$ strands
of the surface $\sg$. We prove that $H_*(\ggm;\Z_2)\cong H_*(\gg;H_*(\bms;\Z_2))$. The main ingredient
is the identification of $H_*(\bms;\Z_2)$ as a symplectic representation of $\gg$.
%We compute (?) the right hand side explicitly for $g=1$.
\end{abstract}

%\thanks{}

\maketitle

\input{Introduction.tex}

\input{Preliminaries.tex}

\input{HBraidSurf.tex}

\input{Actiongg.tex}

\input{Ssequence.tex}
% 
% \input{GenusOne.tex}
% 
\input{Rational.tex}
% 
% \input{Conclusions.tex}

\bibliography{Bibliography.bib}{}
\bibliographystyle{plain}

\end{document}