### 2019-02-28 simpleexample.tex
Drawn and redrawn figures

\subsection{Geometric examples of homology classes}
In this subsection we consider the surface $\Sigma_{2,1}$, which we abbreviate with $\S$, and
construct some homology classes in $H_2(C_2(\S))$: our aim is to get a first understanding
of why this homology group is isomorphic to $\Sym_2(\H)=\H^{\otimes 2}/\mathfrak{S}_2$.