### 2019-07-12 splitting.tex

\documentclass{amsart}
\usepackage{amsmath,amssymb,amsthm,amsfonts,amscd,mathrsfs,mathtools}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage[hyphens]{url}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{tikz-cd}

\setlength\parindent{0pt}

\hyphenation{e-qui-va-riant se-cond ho-mo-lo-gy co-ho-mo-lo-gy}
% \setlength{\parindent}{6pt}

\theoremstyle{plain}
    \newtheorem{thm}{Theorem}[section]
    \newtheorem{lem}[thm]   {Lemma}
    \newtheorem{cor}[thm]   {Corollary}

\theoremstyle{definition}
    \newtheorem{defn}[thm]  {Definition}
    \newtheorem{conj}[thm]{Conjecture}
    \newtheorem{ex}{Example}

\input{macros.tex}

\begin{document}

\title{Splitting of the homology of the punctured mapping class group}

\author{Andrea Bianchi}

\address{Mathematics Institute,
University of Bonn,
Endenicher Allee 60, Bonn,
Germany
}

\email{bianchi@math.uni-bonn.de}

\date{\today}

\keywords{Mapping class group, braid group, symplectic coefficients.}
\subjclass[2010]{20F36, %Braid groups; Artin groups
55N25, %Homology with local coefficients, equivariant cohomology
55R20, %Spectral sequences and homology of fiber spaces 
55R35, %Classifying spaces of groups and $H$-spaces
55R40, %Homology of classifying spaces, characteristic classes
55R80, %Discriminantal varieties, configuration spaces
55T10, %Serre spectral sequences
}

\begin{abstract}
Let $\Gamma_{g,1}^m$ be the mapping class group of the orientable surface $\Sigma_{g,1}^m$ of genus $g$ with one parametrised
boundary curve and $m$ permutable punctures; when $m=0$ we omit it from the notation.
Let $\beta_{m}(\Sigma_{g,1})$ be the braid group on $m$ strands
of the surface $\Sigma_{g,1}$.

\noindent We prove that $H_*(\Gamma_{g,1}^m;\mathbb{Z}_2)\cong H_*(\Gamma_{g,1};H_*(\beta_{m}(\Sigma_{g,1});\mathbb{Z}_2))$. The main ingredient
is the computation of $H_*(\beta_{m}(\Sigma_{g,1});\mathbb{Z}_2)$ as a symplectic representation of $\Gamma_{g,1}$.
%We compute (?) the right hand side explicitly for $g=1$.

% 
% Let $\Gamma^m_{g,1}$ be the mapping class group of an orientable surface $\Sigma^m_{g,1}$ of genus $g$ with one parametrised
% boundary curve and $m$ permutable punctures, and let $\beta_m(\Sigma_{g,1})$ be the braid group on $m$ strands
% of the surface $\Sigma_{g,1}$ of genus $g$ with one boundary curve.
% 
% We prove that $H_*(\Gamma^m_{g,1};\mathbb{Z}_2)\cong H_*(\Gamma_{g,1};H_*(\beta_m(\Sigma_{g,1});\mathbb{Z}_2))$. The main ingredient
% is the computation of $H_*(\beta_m(\Sigma_{g,1});\mathbb{Z}_2)$ as a symplectic representation of $\Gamma_{g,1}$.
\end{abstract}

%\thanks{}

\maketitle
THIS VERSION IS NOT ANYMORE UP TO DATE! SEE ARXIVE VERSION

\input{Introduction.tex}

\input{Preliminaries.tex}

\input{HBraidSurf.tex}

\input{Actiongg.tex}

\input{Ssequence.tex}

\input{Comparison.tex}

% \input{Rational.tex}

\bibliography{Bibliography.bib}{}
\bibliographystyle{plain}

\end{document}