### 2020-05-01 Rational homology/rational_homology.tex

\documentclass{amsart}
\usepackage{amsmath,amssymb,amsthm,amsfonts,amscd,mathrsfs,mathtools}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage[hyphens]{url}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{tikz-cd}

\setlength\parindent{0pt}

\hyphenation{e-qui-va-riant se-cond ho-mo-lo-gy co-ho-mo-lo-gy}
% \setlength{\parindent}{6pt}

\theoremstyle{plain}
    \newtheorem{thm}{Theorem}[section]
    \newtheorem{lem}[thm]   {Lemma}
    \newtheorem{cor}[thm]   {Corollary}

\theoremstyle{definition}
    \newtheorem{defn}[thm]  {Definition}
    \newtheorem{conj}[thm]{Conjecture}
    \newtheorem{ex}{Example}

\input{macros.tex}

\begin{document}

\title{Rational homology of configuration spaces of surfaces}

\author{Andrea Bianchi}

\address{Mathematics Institute,
University of Bonn,
Endenicher Allee 60, Bonn,
Germany
}

\email{bianchi@math.uni-bonn.de}

\date{\today}

\maketitle
Let $C_n(\Sg)$ denote the unordered configuration spaces of $n$ points
on the surface $\Sg$. In this note I list all the properties I know about $H_*\pa{C_n(\Sg);\Q}$.
Many of these properties can be found somewhere in the literature,
From now on rational coefficients in homology are always understood.

Let $C_{\bullet}(\Sg):=\coprod_{n\geq 0}C_n(\Sg)$, which is a module over the 
(homotopy) associative H-space $C_{\bullet}(\C)=\coprod_{n\geq 0}C_n(\C)$.

Then $\M:=H_*(C_{\bullet}(\Sg)$ is a bigraded $\Gamma_{g,1}$-representation over the bigraded Pontryiagin
ring $H_*(C_{\bullet}(\C))\cong\Q[\epsilon]\otimes \Lambda (\alpha)$, where $\epsilon$ has
bidegree (0,1) and $\alpha$ has bidegree (1,2). The bidegree $(*,\bullet)$ is given by
the homological degree $*$ and the weight (number of points used) $\bullet$.

Let $\bH=H_1(\Sg)$, concentrated in bidegree (2,2), and let $\cH=H_1(\Sg)$, concentrated in bidegree (1,1).
The following hold:

\begin{enumerate}
 \item $\M$ is in general neither a free $\Q[\epsilon]\otimes \Lambda (\alpha)$-module (only for $g=0$),
 nor a symplectic $\Gamma_{g,1}$-representation (only for $g\leq 1$, see the \emph{rational counterexample}
 in my article).
 \item Nevertheless $\M$ is a free $\Q[\epsilon]$-module (see work of B\"odigheimer and Tillmann), and the Johnson's kernel
 $\mathcal{J}_{g,1}\subset\Gamma_{g,1}$ acts trivially on it (one can check that
 separating Dehn twists act trivially on a generatring set of $\M$).
 \item Let $\Sym(\bH)$ denote the symmetric algebra on $\bH$, which is a bigraded ring with a
 symplectic action of $\Gamma_{g,1}$. Then $\M$ has a natural structure of (equivariant)
 $\Sym(\bH)$-module, i.e. there is a $\Gamma_{g,1}$-equivariant map
 \[
  \mu\colon\Sym(\bH)\otimes \M\to \M
 \]
 which satisfies the usual associativity, it preserves the bidigree and it is compatible with
 the action of the ring $\Q[\epsilon]\otimes \Lambda (\alpha)$ (this seems true to me, although
 I have never written down a formal proof of it).
 \item One can filter $\M$ by the submodules
 \[
  \M\supset\bH\cdot\M\supset\bH^2\cdot\M\supset\bH^3\cdot \M.
 \]
 Every $\bH^k\cdot\M$ is bigraded, $\Gamma_{g,1}$-invariant, and an equivariant module
 over $\Q[\epsilon]\otimes \Lambda (\alpha)\otimes \Sym(\bH)$. The associated graded,
 denoted $grad\M$, is also a bigraded $\Gamma_{g,1}$-equivariant module over $\Q[\epsilon]\otimes \Lambda (\alpha)\otimes \Sym(\bH)$.
 \item As such, $grad\M$ is isomorphic to the following
 \[
  grad\M\cong \Q[\epsilon]\otimes \Sym(\bH)\otimes [\alpha\otimes\coker(\cdot\wedge\omega)\oplus\Ker(\cdot\wedge\omega)].
 \]
 Here $(\cdot\wedge\omega)\colon\Lambda(\cH)\to\Lambda(\cH)$ is a map of bidegree (-2,-2), given by contracting
 an element of the exterior algebra $\Lambda(\cH)$ with the intersection form of the surface $\omega\in\Lambda_2(\cH^*)$.
 The action of $\Gamma_{g,1}$ is symplectic on $grad\M$. Multiplication by $\alpha$ is trivial on the summand
 $\alpha\otimes\coker(\cdot\wedge\omega)$ and is induced by the natural map
 $\Ker(\cdot\wedge\omega)\subset \Lambda(\cH)\to\coker(\cdot\wedge\omega)$ on the other summand (this should
 also be entailed in the work of Pagaria).
\end{enumerate}
\end{document}